Nuclei Segmentation in Histopathology images using Deep Neural Networks

### 2016-11-07 Paper.tex
Morphology added

\name{Peter Naylor$^{1,2,3}$\sthanks{Peter Naylor has received a PhD
  fellowship from the Ligue contre le Cancer. } \qquad Marick Lae$^{2}$
  \qquad Fabien Reyal$^{2}$\qquad Thomas Walter$^{1,2,3}$}

\address{$^{1}$MINES ParisTech, PSL Research University, CBIO - Centre
  de Bioinformatique,\\ 
  35 rue St Honoré 77300 Fontainebleau, France \\
$^{2}$Institut Curie, 75248 Paris Cedex, France\\
$^{3}$INSERM U900, 75248 Paris Cedex, France}
%
\begin{document}
%\ninept
%
\maketitle
%
\begin{abstract}
%The abstract should appear at the top of the left-hand column of text, about
%0.5 inch (12 mm) below the title area and no more than 3.125 inches (80 mm) in
%length.  Leave a 0.5 inch (12 mm) space between the end of the abstract and the
%beginning of the main text.  The abstract should contain about 100 to 150
%words, and should be identical to the abstract text submitted electronically
%along with the paper cover sheet.  All manuscripts must be in English, printed
%in black ink.
\end{abstract}
%
\begin{keywords}
Deep Learnnig, Convolutional Neural Networks, Nuclei Segmentation,
Histpathology, Digital Pathology, Breast Cancer, Cellular Phenotyping
\end{keywords}
%
\section{Introduction}
\label{sec:intro}

@@

Yet, segmentation of nuclei is a
complicated task: tissue type, staining differences and cell type
convey them different visual characteristics, which makes it very difficult to
design traditional image segmentation algorithms that work
satisfactorily for all of these different cases. On the other hand,
deep learning algorithms have been used recently with great success to
complex segmentation tasks in biology\cite{Ciresan2012,Ronneberger2015}.

@@

\section{Different architectures/Results}
\label{sec:method}
We experience with 4 known arhitectures in semantic segmentation, one 
will be named BaochuanNet, Fully Convolutionnal Net (FCN), DeconvNet, 
UNet. The most basic architecture, BaochuanNet, consists of 4 
convolutionnal layer where each convolutionnal layer has 8 feature map. 
This net, being not deep, has the advantage of being less 
computationnally intensive. FCN is a first attempt of applying "deep 
feature" representations to the task of semantic segmentation. This 
archicture has the advantage of re-using a classical deep learning 
architecture with added on upsampling layer and skip layers. The 
upsampling layers enable the network to learn a pixel level classification 
and the skip layers enable the network to fuse different levels of 
abstraction to the final prediction. This model can be fine tuned with a 
set of pretrained-weights extracted from the classical deep learning 
architecture. DeconvNet is also based on a classical architecture, 
however, in this network their are no skip layer as one intends to learn 
the proper upsampling threw repeated deconvolution and convolution 
layers. This model can also be fine tuned. The final architecture that we tried is the UNet. This network combines both contribution of the above networks, the UNet has a similar architecture to the DeconvNet with skip layers.
\section{Result}
\label{sec:result}

@@

\section{REFERENCES}
\label{sec:ref}

@@

% References should be produced using the bibtex program from suitable
% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
% style file from IEEE produces unsorted bibliography list.
% -------------------------------------------------------------------------
\bibliographystyle{IEEEbib}
\bibliography{strings}

>>>>

\name{Peter Naylor$^{1,2,3}$\sthanks{Peter Naylor has received a PhD
  fellowship from the Ligue contre le Cancer. } \qquad Marick La\'e$^{4}$
  \qquad Fabien Reyal$^{2}$\qquad Thomas Walter$^{1,2,3}$}

\address{$^{1}$MINES ParisTech, PSL Research University, CBIO - Centre
  de Bioinformatique,\\ 
  35 rue St Honoré 77300 Fontainebleau, France \\
$^{2}$Institut Curie, 75248 Paris Cedex, France\\
$^{3}$INSERM U900, 75248 Paris Cedex, France\\
$^{4}$Service de Pathologie, Institut Curie, 75248 Paris Cedex, France}
%
\begin{document}
%\ninept
%
\maketitle
%
\begin{abstract}
%The abstract should appear at the top of the left-hand column of text, about
%0.5 inch (12 mm) below the title area and no more than 3.125 inches (80 mm) in
%length.  Leave a 0.5 inch (12 mm) space between the end of the abstract and the
%beginning of the main text.  The abstract should contain about 100 to 150
%words, and should be identical to the abstract text submitted electronically
%along with the paper cover sheet.  All manuscripts must be in English, printed
%in black ink.
\end{abstract}
%
\begin{keywords}
Deep Learnnig, Convolutional Neural Networks, Nuclei Segmentation,
Histpathology, Digital Pathology, Breast Cancer, Cellular Phenotyping
\end{keywords}
%
\section{Introduction}
\label{sec:intro}

@@

Yet, segmentation of nuclei is a
complicated task: tissue type, staining differences and cell type
convey them different visual characteristics, which makes it very difficult to
design traditional image segmentation algorithms that work
satisfactorily for all of these different cases. On the other hand,
deep learning algorithms have been used recently with great success to
complex segmentation tasks in biology\cite{Ciresan2012,UNet}.

@@

%% Mathematical Morphology
The results obtained by the Deep Neural Network were encouraging, but
touching nuclei were often segmented as one single object. This is not
surprising, as the error at a pixel level seems very minor, and no
shape prior is used in the model. Solving this issue by weighting the
error term for pixels between objects did not solve the issue for our
case (data not shown). However, we did observe that for touching and
even partially overlapping nuclei, the posterior probability at the
nucleus border is systematically lower than in
the putative center of the nucleus, but may still be larger than
typically applied thresholds.This leads to the following rationale. As in the center of nuclei, the
posterior probability is maximal, we can readily assume that the local
maxima of the posterior correspond to putative nuclei, which we call candidates. Let
$\mathcal{X}=(x_1,x_2,\ldots,x_N)$be a path that joins two candidates
(i.e. $\forall i x_i$ and $x_{i+1}$ are neighbor pixels) and
$\mathcal{P}=(p_1,p_2,\ldots,p_N)$ the corresponding posterior
probabilities. Without loss of generality, we assume $p_1\leq P_N$.
We define for each path $\mathcal{X}$ a cost
$C(\mathcal{P})=\max_{i=2 \ldots N}{p_1-p_i}$, which is the maximal
decrease in posterior probability along the path, when starting from
the candidate with lower probability.  Considering all paths joining two
candidates, we can now state a criterion that allows us to decide whether to
perform a split (and thus accept the candidates as being different
nuclei) or not: if all the paths involve a decrease in probability
larger than a parameter $\lambda$, we will perform a
split. Conversely, if we can find a path that joins the two candidates
with a probability decrease smaller or equal to $\lambda$, no split
will be performed: we argue that in this case, it is probably one
single object. Hence, the split is performed if: 
\begin{equation}
\min_{\mathcal{P}}C(\mathcal{P}) = \min_{\mathcal{P}} \{\max_{i=2
  \ldots N}{p_1-p_i}\} > \lambda
\end{equation}
where $\lambda$ is a free parameter. This is actually nothing else
than the morphological dynamics\cite{Grimaud1992}. The actual split
locations are thus obtained by applying the watershed algorithm,
starting from the local maxima of the posterior probability map with
morphological dynamics larger than a free parameter $\lambda$.

To speed up this procedure, we have developed an algorithm that
performs selection and Watershed algorithm in a single pass by
selectively building the Watershed line or fusing the regions,
depending on the morphological dynamic at the moment at which two
regions get in contact. This algorithm will be made publicly available
through the open-source software scikit-image\cite{scikit-image}, in which we implemented
all the steps which were not related to deep learning.

\section{Different architectures/Results}
\label{sec:results}
We experience with 4 known arhitectures in semantic segmentation, one 
will be named BaochuanNet, Fully Convolutionnal Net (FCN), DeconvNet, 
UNet. The most basic architecture, BaochuanNet, consists of 4 
convolutionnal layer where each convolutionnal layer has 8 feature map. 
This net, being not deep, has the advantage of being less 
computationnally intensive. FCN is a first attempt of applying "deep 
feature" representations to the task of semantic segmentation. This 
archicture has the advantage of re-using a classical deep learning 
architecture with added on upsampling layer and skip layers. The 
upsampling layers enable the network to learn a pixel level classification 
and the skip layers enable the network to fuse different levels of 
abstraction to the final prediction. This model can be fine tuned with a 
set of pretrained-weights extracted from the classical deep learning 
architecture. DeconvNet is also based on a classical architecture, 
however, in this network their are no skip layer as one intends to learn 
the proper upsampling threw repeated deconvolution and convolution 
layers. This model can also be fine tuned. The final architecture that we tried is the UNet. This network combines both contribution of the above networks, the UNet has a similar architecture to the DeconvNet with skip layers.
\section{Result}
\label{sec:result}

@@

%\section{REFERENCES}
%\label{sec:ref}

@@

% References should be produced using the bibtex program from suitable
% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
% style file from IEEE produces unsorted bibliography list.
% -------------------------------------------------------------------------
\bibliographystyle{IEEEbib}
\bibliography{refs}
